Stochastic Variance Reduced Policy Gradient

### 2018-02-02 svrpg.tex

\section{Hyperparameters}\label{sec:stopping}
In Reinforcement learning we have to take into account the gradient's variance.
\todomat{$\uparrow$ the opening of the section could connect better with previous sections. Declare which meta-parameters will be addressed in the section}
A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter. Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.
We use two different adam hyperparameters: one for updates tied to snapshot points and the other tied to the updates of sub-iterations points.
\todomat{$\uparrow$ explain better: it's not only two separate meta-parameters, but two separate learning rate schedules. Use notation from previous sections, \eg snaphot is usually associated with $\tilde{\vtheta}$} 
This is because we are dealing with two different variance magnitudes.
\todomat{$\uparrow$ the reason is this, but try to spend more words on it}
Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.

As shown in Section \ref{sec:conv}, the performance of the SVRPG algorithm is sensitive to the epoch size $m$.
As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
A fixed epoch size is not the optimal choice.
Few sub-iterations lead to few updates, many update lead to increases of variance.

In order to cope with this behaviour we introduce an adaptive stopping condition: we perform a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot. This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
\todomat{I find "two stopping conditions" confusing: it's just one with an additional constraint}
\todobep{I tried to solve the issue here:hope it's better now}
We stop with inner iterations if:

@@

\todomat{why? should mention how this formula combines trajectory efficiency and variance control}
After that we sample N trajectories for the next Snapshot point.

\section{Practical Variants}\label{sec:prac}
\todomat{Missing opening to section}
As reported in Section \ref{subsec:PolicyGradient} we use the G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance. The off-policy gradient estimator becomes:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
\end{align*} 
where:
\begin{align*}
\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
 }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\begin{align*}
\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
\section{Related Work}

>>>>

\section{Hyperparameters}\label{sec:stopping}
In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot. The first one is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots. The second one is essetial to control the variance introduced by the importance weights, mainly because the farther you go from the snapshot policy the higher the variance of the importance weights will be.

A common way to reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter. Adam helps to stabilize the gradient estimate by computing adaptive learning rates for each parameter based on the history of the gradients' variance.

Due to this feature we can use two different adam in order to properly leverage the structure of the proposed SVRPG update. The first adam will be associated to the snapshot updates and the second one will be associated to the sub-iteration updates. By doing so we are decoupling the contribution to the variance introduced by the snapshots from that one introduced by the sub-iterations. These two variance contributions lie, respectively, on two different order of magnitude because the number of trajectories used to perform the estimates for the snapshot is greater than that used for a sub-iteration.Therefore, using two adam optimizers allows us to properly adapt the learning rate in each situation. For the above mentioned reasons let us denote by $\alpha_{snap}$ the learning rate associated to the last snapshot and by  $\alpha_{sub-iter}$ the learning rate associated to the current sub-iteration.

We have already said that as the number of sub-iterations $m$ increases the variance of the importance weights increases too. Therefore, if $m$ is not properly set the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.

We have seen, sperimentally, that a fixed epoch size is not the optimal choice, but an adaptive one would better exploit the total amount of data at hand. Hence, we permorm a new snapshot anytime the weighted learning rate proposed by the adam optimizer in a sub-iteration is smaller than the weighted one proposed at the immediately previous snapshot. This stopping condition is subject to a constraint on the maximum number of sub-iterations performable, which is problem dependent.
More precisely we stop performing sub-iterations when:

@@

These two weights ${\bf{\it N}}$ and ${\bf{\it B}}$ represent, respectively, the number of trajectories used for a snapshot and the number of trajectories used for an update of the parameters in a sub-iteration. Therefore, in this stopping condition, we use them to take into account the trajectories efficiency. The less trajectories we use to perform an update, the better it is from a data usage point of view. Hence we want to facilitate sub-iterations with respect to snapshots.

Another insight about this adaptive stopping condition lies within the value of the two different learning rate proposed by adam. Whenever the learning rate of a sub-iteration becomes too small with respect to that one proposed at the previous snapshot we are experiencing high variance. Hence it could be better to perform a new snapshot so to mitigate this issue.

\section{Practical Variants}
\todomat{Missing opening to section}
We introduce a term with importance weights to guarantee proper sampling of trajectories.
This term's variance is subjected to the variance of importance weights. 
There are some methods to reduce this variance.

As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance. The off-policy gradient estimator becomes:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
\end{align*} 
where:
\begin{align*}
\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
 }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\begin{align*}
\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
\section{Related Work}
